The Astropy Project

### 2017-10-27 main.tex
Update on Overleaf.

% Packages / projects / programming - for consistency!
\newcommand{\package}[1]{\texttt{#1}\xspace}
\newcommand{\github}{\package{GitHub}\xspace}
\newcommand{\python}{\package{Python}\xspace}
\newcommand{\astropy}{Astropy\xspace}
\newcommand{\astropypkg}{\package{astropy}\xspace}

@@

The first public release of the \astropy package is described in
\cite{astropy}. Since then, the \astropy package has been
used in hundreds of projects, and the scope of the package has grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropy core has
developed. In this paper, we describe the current status of the
\astropy community, the \astropy core package, and the \astropy\
ecosystem.

@@

\subsection{Concept of affiliated packages}
% it may be described in previous subsection, if not the Brigitta can try to
% write it up
%
\par A major part of the \astropy~project is the concept of
``Affiliated Packages''. An affiliated package is an astronomy-related
\python~package that is not part of the \astropypkg~core package, but
has requested to be included as part of the \astropy project's
community. These packages support the goals and vision of \astropy~of
improving code reuse, interoperability, and embracing good coding
practices such as testing and thorough documentation. \astropypkg{}
version 2.0 supports \python~2.7 and \python~3.4+, affiliated packages
are strongly encouraged to support at least \python~3.x.
%
\par Affiliated packages contain functionality that is more specialized, or
have license incompatibility, or have enough external dependencies (e.g. GUI
libraries) that limit their integration possibility into the core and thus
they are more suitable to be separate packages. Starting as an affiliate 
package is also seen as a good strategy for functionality that is aimed for
the \astropypkg~core but is initially undergoing rapid development and can 
benefit from a different release cycle while stabilizing its API. 
%
\par In general we hope that becoming an affiliated package is seen as 
a good way for new and existing packages to gain exposure.
%
\inlinecomment{BMS}{TODO: write about the process of becoming affiliated, and project packages into the future section below}
%
%\subsection{Mixin}

@@

The \package{astropy.nddata} supackage provides three types of functionality:
an abstract interface for representing generic arbitrary-dimensional datasets
intended primarily for developers of other packages and concrete classes
building on this interface and utilities for manipulating these kind of
datasets.

@@

The \package{astropy.nddata.utils} module provides utilities that can operate on either plain NumPy arrays or any of the classes in the \package{astropy.nddata} subpackage.  It features a class for representing two-dimensional image cutouts, allowing one to easily link pixels in the cutout to pixels in the original image or from the image to the cutout, and to overlay the cutout on images. Functions to enlarge or reduce an image by doing block replication or reduction are also provided.

@@

@@

\package{astropy.modeling} provides several fitters which are wrappers around some of the \texttt{numpy} and \texttt{scipy.optimize} functions and provide supports for constraints. They take a model as input and return a copy of the model with the fitted parameters. The goal is to make it easy to extend the fitting framework and create new fitters. The optimizers available in astropy 2.0 are Levenberg-Marquardt, Simplex, SLSQP, and LinearLSQFitter (which is based on \texttt{numpy.linalg} that provides exact solution for linear models).

Modeling also supports a plugin system for fitters, which allows using the
astropy models with external fitters. An example of this is
SABA\footnote{\url{https://github.com/astropy/saba}}, which is a bridge between
Sherpa\footnote{\url{http://cxc.cfa.harvard.edu/contrib/sherpa/}}
and astropy.modeling, to bring the Sherpa fitters into astropy.

@@

\section{Infrastructure for affiliated packages}
\label{sec:infrastructure}
%\inlinecomment{BMS}{should we use links to the appropriate github repo?}
%\inlinecomment{PLL}{IMHO a list of GitHub repo links for affil. pkgs. should be in appendix, if included.}
\inlinecomment{BMS}{should we describe astropy-helpers even though it's future is unsure?}
%
\par In addition to astronomy related packages and libraries, the \astropy\
Project also maintains and distributes several infrastructure packages that are
very generic. The following sections describes the most widely used
selection of them.
%
\subsection{Package template}
% Brigitta may write this up if there is no other takers
\par \astropy provides a \package{package-template}\footnote{\url{https://github.com/astropy/package-template/}}
package that any \python package is welcome to
use (many affiliated packages do so). The template contains a ready-to-go
package layout. It provides infrastructure such as documentation tools,
testing framework and CI templates and configurations, Cython integration,
and documentation on how to make it all work.
%
\subsection{Continuous integration helpers}
% Brigitta
\package{ci-helpers}\footnote{\url{https://github.com/astropy/ci-helpers}} is a
collection of scripts that empowers package maintainers to control their testing
set up and installation process for
various continuous integration services via a set of environment
variables. While the current development is mostly driven by the needs of
the \astropy ecosystem, the actual usage of this package is extremely
widespread. Currently we support setups for Travis CI and Appveyor CI.
%
\subsection{Sphinx extensions}

@@

dropping \python 2 support, growths of affiliated packages
Summary
learning astropy

@@

At the same time, \astropy is spreading awareness of best practices in
software development. This is important since most practicing
astronomers received their education without being taught in computer
science and software development although a major fraction of almost
every astronomer's workload today is related to software use and
development. The \astropypkg package leads by example, showing all interested
astronomers how modern tools like git version control or continuous
integration (code is tested after every change to the source) can
increase the quality and accessibility of astronomical software
without overly complicating the development cycle. All submitted code
is reviewed by at least one, but typically more, members of the
astropy community, providing feedback and allowing contributors to
improve their skill is developing software without the need for their
institutions to provide dedicated teaching resources. Thus, the
\astropy project does not only build the tools needed to perform
productive astronomical research in an era of increasingly large and
complex datasets, it also helps to prepare the current and the next
generation of researchers with the knowledge to adequately use those
tools.

>>>>

% Packages / projects / programming - for consistency!
\newcommand{\package}[1]{\texttt{#1}\xspace}
\newcommand{\github}{\package{GitHub}}
\newcommand{\python}{\package{Python}}
\newcommand{\astropy}{Astropy\xspace}
\newcommand{\astropypkg}{\package{astropy}}

@@

The first public release of the \astropy package is described in
\cite{astropy}. Since then, the \astropy package has been
used in hundreds of projects, and the scope of the package has grown
considerably. At the same time, the community of astronomers
contributing to the project has grown tremendously and an ecosystem
of packages supporting or affiliated with the \astropy core has
developed. In this paper, we describe the current status of the
\astropy community, the \astropypkg core package, and the \astropy
ecosystem.

@@

\subsection{Concept of affiliated packages}
% it may be described in previous subsection, if not the Brigitta can try to
% write it up
%
\par A major part of the \astropy project is the concept of
``Affiliated Packages''. An affiliated package is an astronomy-related
\python~package that is not part of the \astropypkg core package, but
has requested to be included as part of the \astropy project's
community. These packages support the goals and vision of \astropy of
improving code reuse, interoperability, and embracing good coding
practices such as testing and thorough documentation. \astropypkg{}
version 2.0 supports \python~2.7 and \python~3.4+, affiliated packages
are strongly encouraged to support at least \python~3.x.
%
\par Affiliated packages contain functionality that is more specialized, or
have license incompatibility, or have enough external dependencies (e.g. GUI
libraries) that limit their integration possibility into the core and thus
they are more suitable to be separate packages. Starting as an affiliate 
package is also seen as a good strategy for functionality that is aimed for
the \astropypkg~core but is initially undergoing rapid development and can 
benefit from a different release cycle while stabilizing its API. 
%
\par In general we hope that becoming an affiliated package is seen as 
a good way for new and existing packages to gain exposure.
%
\inlinecomment{BMS}{TODO: write about the process of becoming affiliated, and project packages into the future section below}
%
%\subsection{Mixin}

@@

The \package{astropy.nddata} supackage provides three types of functionality:
an abstract interface for representing generic arbitrary-dimensional datasets
intended primarily for developers of other packages, concrete classes
building on this interface and utilities for manipulating these kind of
datasets.

@@

The \package{astropy.nddata.utils} module provides utilities that can operate on either plain NumPy arrays or any of the classes in the \package{astropy.nddata} subpackage.  It features a class for representing two-dimensional image cutouts, allowing one to easily link pixels in the cutout to pixels in the original image or from the image to the cutout, to convert between world and pixel coordinates in the cutout and to overlay the cutout on images. Functions to enlarge or reduce an image by doing block replication or reduction are also provided.

@@

\inlinecomment{Yes, done.}

@@

\package{astropy.modeling} provides several fitters which are wrappers around some of the \texttt{numpy} and \texttt{scipy.optimize} functions and provide supports for constraints. They take a model as input and return a copy of the model with the fitted parameters. The goal is to make it easy to extend the fitting framework and create new fitters. The optimizers available in \astropypkg 2.0 are Levenberg-Marquardt, Simplex, SLSQP, and LinearLSQFitter (which is based on \texttt{numpy.linalg} that provides exact solution for linear models).

Modeling also supports a plugin system for fitters, which allows using the
astropy models with external fitters. An example of this is
\package{SABA}\footnote{\url{https://github.com/astropy/saba}}, which is a bridge between
Sherpa\footnote{\url{http://cxc.cfa.harvard.edu/contrib/sherpa/}}
and \package{astropy.modeling}, to bring the Sherpa fitters into astropy.

@@

\section{Infrastructure for affiliated packages}
\label{sec:infrastructure}
%\inlinecomment{BMS}{should we use links to the appropriate github repo?}
%\inlinecomment{PLL}{IMHO a list of GitHub repo links for affil. pkgs. should be in appendix, if included.}
\inlinecomment{BMS}{should we describe astropy-helpers even though it's future is unsure?}
%
\par In addition to astronomy related packages and libraries, the \astropy
Project also maintains and distributes several generic purpose infrastructure packages.
The following sections describes the most widely used selection of them.
%
\subsection{Package template}
% Brigitta may write this up if there is no other takers
%
\par \astropy provides the \package{package-template}\footnote{\url{https://github.com/astropy/package-template/}}
package. Using this template, any \python package can make use of the ready-to-go package
layout, setup, installation and documentation infrastructure developed originally for the
\astropypkg core package and affiliated packages maintained by the \astropy. It also provides
a testing framework, CI templates and configurations, Cython integration, and documentation
on how to make it all work.
%
\subsection{Continuous integration helpers}
% Brigitta
%
\package{ci-helpers}\footnote{\url{https://github.com/astropy/ci-helpers}} is a
collection of scripts that empowers package maintainers to control their testing
set up and installation process for
various continuous integration services via a set of environment
variables. While the current development is mostly driven by the needs of
the \astropy ecosystem, the actual usage of this package is extremely
widespread. Currently we support setups for Travis CI and Appveyor CI.
%
\subsection{Sphinx extensions}

@@

dropping \python{} 2 support, growths of affiliated packages

\subsection{Learning astropy}

@@

At the same time, \astropy is spreading awareness of best practices in
software development. This is important since most practicing
astronomers received their education without being taught in computer
science and software development although a major fraction of almost
every astronomer's workload today is related to software use and
development. The \astropypkg package leads by example, showing all interested
astronomers how modern tools like git version control or continuous
integration (code is tested after every change to the source) can
increase the quality and accessibility of astronomical software
without overly complicating the development cycle. All submitted code
is reviewed by at least one, but typically more, members of the
\astropy community, providing feedback and allowing contributors to
improve their skill is developing software without the need for their
institutions to provide dedicated teaching resources. Thus, the
\astropy project does not only build the tools needed to perform
productive astronomical research in an era of increasingly large and
complex datasets, it also helps to prepare the current and the next
generation of researchers with the knowledge to adequately use those
tools.
